Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Update intro
Update background

\section{Introduction}

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.49\textwidth]{./figs/dish_overview.pdf}
\caption{
  \textbf{Applying \sys to $p_1$.}
}
\label{fig:example}
\end{figure*}

@@

Pipelines are a mechanism for program composition.
Individual programs, specified as commands, conform to a common but general interface. 
For now, this interface allows a combination of the following:
  accepting input from the standard input stream (stdin),
  generating output to the standard output stream (stdout), and
  generating errors or diagnostics to the standard error stream (stderr)
Standard streams are streams of characters with a few special characters---for example, EOF is a special end-of-file character signaling the end of the input stream.
% For Unix commands, the newline character is special; as such, Unix streams can be seen as lazy lists of strings---lines of text separated by the newline character.

Leveraging this interface, pipelines chain together commands by their standard streams, such that the output stream of one command (stdout) is passed directly as input (stdin) to the next one.
Pipeline chaining goes hand-in-hand with the Unix toolbox philosophy: 
  programs do one thing well, they work together, and they manipulate text streams as their universal interface.
The next few paragraphs describe the details.

@@

\heading{Generation}
A central concept in pipelines---and Unix, in general---is the file system, which allows interfacing with persistent data.
There are many commands that take as input a file identifier 
Reading a file or a set of files generates a stream; 
  similarly, streams can be trivially redirected to the file system---by appending with the file identifier.

Expansion---the key is that expansion and evaluation are handled by the shell at an earlier stage 
The most powerful---and interesting, for \ttt{dish}---type of expansion is subshell expansion, that allows replacing entire input streams

@@

\kk{We have to mention character vs line based parallelism somewhere,
  and that we choose to see lines as the data quantum}

The file-system is a common abstraction layer used for many other operations than just serving files (this will be become clearer when talking about file-system commands in \S\ref{bg:cmd}).
as such, different identifiers many not resolve to actual files, but rather in-memory or OS-internal structures.

At a high level, POSIX defines several file types:
(-) regular files, storing lines of text or binary data,
(d) directories, grouping together multiple other files,
(l) symbolic links, pointing to other files,
(p) named pipes or FIFOs, communication primitives that can be named and manipulated like files,
(s) domain sockets or DSes, full-duplex communication primitives that support passing file descriptors, and
(b or c) special files, interfaces to device drivers or low-level abstractions, presented as ordinary files.

Most of these identifiers are manipulated by file-system commands---\eg \ttt{mkdir} for directories and \ttt{mkfifo} for FIFOs.
FIFOs and DSes look like files but are streams.

Special files are particularly interesting because they are resolved to OS structures.
One set is pseudo-devices---device files that simply;
For example /dev/urandom returns a stream of random characters and /dev/null accepts and discards all input written to it.
Another set is proc filesystem (procfs), a special filesystem in Unix-like operating systems that presents information about processes and other system information.
For example, /proc/PID/cmdline contains command that originally started the process and /proc/PID/cwd contains a symlink to the current working directory of the process.

sysfs is a pseudo file system provided by the Linux kernel that exports information about various kernel subsystems, hardware devices, and

/modules, one of the most important files in /proc, containing a list of the kernel modules currently loaded . It gives some indication (not always entirely correct) of dependencies.
/cpuinfo

symbolic links, FIFO specials, 
device files
block special, character special, 
and domain sockets.

@@

\section{Related Work}

>>>>

\section{Introduction}
\label{intro}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{./figs/dish_schematic.pdf}
\caption{
  \textbf{\sys high-level schematic.}
	\sys XYZ
}
\label{fig:schematic}
\end{figure}

@@

% \begin{figure*}[t]
% \centering
% \includegraphics[width=0.49\textwidth]{./figs/dish_overview.pdf}
% \caption{
%   \textbf{Applying \sys to $p_1$.}
% }
% \label{fig:example}
% \end{figure*}
%

@@

A pipeline is a mechanism for program composition that works by chaining together programs (or, \emph{commands}) using pipes.
Behind the scenes, \unix facilitates program scheduling and communication synchronization.
Intermediate programs run concurrently consuming input and producing output;
  they can be developed in any programming language, as long as they conform to a common interface.
This interface allows any combination of the following:
  accepting input from the standard input stream (stdin),
  generating output to the standard output stream (stdout), and
  generating errors or diagnostics to the standard error stream (stderr)
These standard streams are lines of characters, with a few special characters---for example, EOF is a special end-of-file character signaling the end of a stream.
Leveraging this interface, pipelines chain together commands by their standard streams, such that the output stream of one command (stdout) is passed directly as input (stdin) to the next one.
% TODO: \kk{We have to mention character vs line based parallelism somewhere, and that we choose to see lines as the data quantum}

Along with their implicit context---these standard streams and environment variables---commands can express
Command flags (options)

% TODO: Talk about security?
Streams are a key feature of \unix, and go hand-in-hand with the file abstraction:
  files are simply persistent streams that can be manipulated by commands or pipelines.
Reading a file or a set of files generates a stream; 
  similarly, streams can be trivially redirected to the file system---by appending with the file identifier.
Files are named, in the sense that they correspond to a global identifier---for example \ttt{cat /x/y} returns the persistent stream associated with the identifier \ttt{/x/y}.
\unix exposes a few different types of files---\eg directories, symbolic links, named pipes (or FIFOs), and special character devices.
As a result, file-system identifiers part of a pipeline may resolve to in-memory structures or pseudo-devices (\eg \ttt{/dev/urandom}).

% At a high level, POSIX defines several file types:
% (-) regular files, storing lines of text or binary data,
% (d) directories, grouping together multiple other files,
% (l) symbolic links, pointing to other files,
% (p) named pipes or FIFOs, communication primitives that can be named and manipulated like files,
% (s) domain sockets or DSes, full-duplex communication primitives that support passing file descriptors, and
% (b or c) special files, interfaces to device drivers or low-level abstractions, presented as ordinary files.

Special files are particularly interesting because they are resolved to OS structures.
One example is the proc filesystem (procfs), a special file-system that presents information about processes and other system information.
This file-system enables a form of programmatic introspection, in which commands can inspect their directory or the flags passed to them.
% Here are a few examples:
% (i) \ttt{/proc/PID/cmdline} contains the command that originally started the process,
% (ii) \ttt{/proc/modules} contains a list of the kernel modules currently loaded, and
% (iii) \ttt{/proc/PID/cwd} contains a symlink to the current working directory of the process.
Similarly, \unix exposes file-system handles for various process-related information, include identifiers for ephmeral streams---\ie ones 
OS-internal structures.

This isomorphism is bidirectional---

---and a key enabler of the runtime rebinding achieved during our transformations.
between file-system handles and ephmeral streams

% For Unix commands, the newline character is special; as such, Unix streams can be seen as lazy lists of strings---lines of text separated by the newline character.

% Pipeline chaining goes hand-in-hand with the Unix toolbox philosophy: 
%   programs do one thing well, they work together, and they manipulate text streams as their universal interface.

\begin{figure}[t]
\centering
\includegraphics[width=0.49\textwidth]{./figs/dish_overview.pdf}
\caption{
  \textbf{\sys transformation overview.}
	\sys XYZ
}
\label{fig:overview}
\end{figure}

@@

\heading{Generation}
Expansion---the key is that expansion and evaluation are handled by the shell at an earlier stage 
The most powerful---and interesting, for \ttt{dish}---type of expansion is subshell expansion, that allows replacing entire input streams

@@

@@

\section{Related Work}
\label{related}

\section{Conclusion}
\label{discussion}
